### 2017-02-20 paper/motivation.tex

\section{Motivation}

In this section, we motivate mergeable data types and \name
programming model via a series of examples.

\subsection{Counter}

\begin{figure}

\begin{subfigure}[b]{0.4\textwidth}
  \begin{ocaml}
    module Counter: sig
      type t
      val add: int -> t -> t
      val mult: int -> t -> t
      val read: t -> int
    end = struct
      type t = int
      let add x v = v + (abs x)
      let mult x v = v * (abs x)
      let read v = v
    end
  \end{ocaml}

\caption{\C{Counter} library in OCaml}
\label{fig:counter-adt}
\end{subfigure}
\begin{subfigure}[b]{0.56\textwidth}
  \begin{ocaml}
    module Counter: sig
      type t
      type eff = t -> t
      val add: int -> t-> eff
      val mult: int -> t -> eff
      val read: t -> int
    end = struct
      type t = int
      type eff = t -> t
      let add x v = fun v' -> v' + (abs x)
      let mult x v = 
        fun v' -> v' + v * (abs x - 1) 
      let read v = v
    end
  \end{ocaml}

\caption{\C{Counter} library re-engineered for op-based replication}
\label{fig:counter-rdt}
\end{subfigure}
% \begin{minipage}{0.62\textwidth}
%   \begin{ocaml}
%     module Counter = struct
%       include Counter
%       type t = counter [@@deriving persistence]
%       let merge old v1 v2 = old + (v1 - old) 
%                                 + (v2 - old)
%     end
%   \end{ocaml}
% \end{minipage}

% \caption{\C{Counter} library in OCaml equipped with a \C{merge}
% function. The \C{deriving} syntax~\cite{ppx-deriving} is processed by
% \name's meta-programming library to automatically derive a persistent
% variant of the \C{Counter}}

% \label{fig:counter}
\end{figure}

We will start with a simple example of a counter that nonetheless
demonstrates the benefits of our approach. An OCaml implementation of
the \C{Counter} library is shown in Fig.~\ref{fig:counter-adt}.
\C{Counter} supports two (update) operations - \C{add} and \C{mult} -
that let a non-negative integer value to be added to the counter, and
multiply the counter, respectively. Observe that the library is
written in an idiomatic style in OCaml; any OCaml programmer faced
with the task of implementing the counting functionality would
presumably write the same.

The counter implementation of Fig.~\ref{fig:counter-adt} serves the
application well as long as it is used in-memory on a single machine.
However, as the application grows in complexity and scales beyond a
single machine, its state may also need to be replicated and
persisted. Unfortunately, replication doesn't come for free, and
existing techniques require significant re-engineering of the
application and its libraries to support replication. A method often
adopted~\cite{crdts,gotsman-popl16} is to re-define the operations to
return \emph{effects} instead of values.  For instance, a \C{Counter}
operation returns a function of type $\C{Counter.t} \rightarrow
\C{Counter.t}$, which is then applied to the counter values at all
replicas to realize its effect.  Fig.~\ref{fig:counter-rdt} shows the
\C{Counter} library with operations re-defined to returns effects.
Such re-engineering has to be done carefully, lest an operation may
may end up causing an unintended effect at a remote replica. For
instance, while the effect of a \C{mult x v} operation may intuitively
seem to be \C{fun v' -> (abs x) * v'}, it is incorrect for it affects
different replicas differently, depending on the local counter value.
Fig.~\ref{fig:counter-rdt} therefore defines the result of \C{mult x
v} as an effect that increments the counter value by \C{v * (abs x -
1)} uniformly at all replicas. Likewise, accidentally defining the
effect of an \C{add x v} operation as \C{fun v' -> (abs x) + v}
results in replacing the counter value at a remote replica with the
one at origin replica, thereby losing updates at the remote replica.
Thus, reasoning about the correctness of effects often involves
appeals to system-specific artifacts, such as replicas, which don't
normally manifest in the program.  While checking the effects for
commutativity may help catch some of these oversights, commutativity
is a non-local property, and automatic verification is difficult in
general. Besides the possibility of new bugs, re-engineering libraries
to return effects has obvious software engineering problems in the
sense that it breaks backwards compatibility with the existing code.