### 2017-02-25 paper/intro.tex
edits

\section{Introduction}
\label{sec:intro}

Real-world distributed programs are challenging to write and maintain
because they often conflate two distinct mechanisms.  The first
concerns the expression of application logic - how do we define
computations that are robust in the presence of distributed
communication among concurrently executing threads of control?  The
second deals with system concerns - how do we express notions of
visibility, replication, and consistency when the nodes participating
in such computations may be geographically distributed and the
networks that connect them unreliable?  To simplify reasoning in such
complicated environments, programming models often make strong
assumptions on the guarantees provided by an implementation (such as
serializability~\cite{Serializability} or strong
consistency~\cite{CDE+12,DNN+15}) that may inadvertently mask,
restrict, or ignore important albeit unpleasant realities (e.g.,
network partitions~\cite{Brewer2000,Gilbert2002}). Given these
unwarranted assumptions, program behavior is often difficult to
predict and verify.  On the other hand, when these features are
explicitly exposed to the programmer, e.g., by requiring that
applications be written in terms of specialized distributed data
structures~\cite{Burckhardt2014,BFL+12,SPB+11} or control
primitives~\cite{Calm}, simplicity, composability, and
ease-of-reasoning can suffer.

To illustrate how these kinds of problems manifest in practice,
consider how we might write a simple counter library (see
Fig.~\ref{fig:counter-adt}).  A \C{Counter} supports two (update)
operations - \C{add} and \C{mult} - that lets a non-negative integer
value be added or multiplied to the counter, resp.  Observe that the
library is written in an idiomatic functional style, with no special
reasoning principles needed to realize desired functionality.  As long
as applications use the library on a single machine, this
implementation behaves as expected.  However, if the library is used
in the context of a more sophisticated application, say one whose
computation is distributed among a collection of machines, its
behavior can become significantly harder to understand.  In
particular, a distributed implementation might wish to
\emph{replicate} the counter state on each node to improve response
time or fault tolerance.  Unfortunately, adding replication doesn't
come for free.  Attempting to update every replicated copy atomically
is problematic in the absence of sophisticated transaction support,
which impose significant performance penalties.  But, without such
heavyweight mechanisms, applying an \C{Add} operation on one node may
not be instanteously witnessed on another, which may in the process of
simultaneously attempting to perform its own \C{Add} or \C{Mult}
action.  While synchronizing the activities of all nodes to ensure at
most one such operation is performed at a time is impractical,
designating a single node to hold the counter state eschewing
replication altogether (as in a client-server
configuration~\cite{Armstrong}), is also an undesirable solution,
given its sensitivity to network partitions and server failures, and
the negative performance impact it incurs in geo-distributed
environments~\cite{Walter}.  Removig coordination altogether by
simply replicating counter state without having any supporting
consistency protocol is an equally infeasible approach.

A method often adopted~\cite{SPB+11,Burckhardt2014} to address this
tension is to re-define a datatype's operations to return
\emph{effects} instead of values.  An \emph{effect} is a tag that
identifies the operation to be applied uniformly at all replicas to
incorporate the effects of the original
operation. Fig.~\ref{fig:counter-rdt} shows the \C{Counter} library
with operations re-defined to returns effects.  Observe that the
\C{Counter.add x v} operation now returns an \C{Add x} effect, which,
when applied at a replica (see \C{apply} in the figure), adds \C{x} to
the local counter value.  Note, however, that \C{add} and \C{mult} are
not commutative operations - assuming two replicas have the same
initial counter value of 0, applying the effect of adding 3 and then
multiplying 5 on one replica yields 15, while applying the effect of
first multiplying 5 and then adding 3 yields 3 on the other.  Thus,
implementations must be carefully written to take the lack of
commutativity into account when defining how effects are applied at a
replica; in the figure, multiplication is expressed in terms of
addition to avoid the kind of undesirable behavior described above.

In this implementation, all replicas will eventually contain the same
counter value, assuming computation on the counter quiesces.  While
transmitting effects provide a low-level operational basis for
handling replicated state, the \emph{ad hoc} nature of the solution
confounds desirable high-level reasoning principles.  Indeed, the
semantic gap between the two versions of the counter, one cognizant of
replication and the other not, breaks backward compatibility.  Just as
significantly, its non-trivial construction must be developed in
different guises for every distributed data structure used by the
application.
\begin{figure}
\begin{subfigure}[b]{0.4\textwidth}
  \begin{ocaml}
    module Counter: sig
      type t
      val add: int -> t -> t
      val mult: int -> t -> t
      val read: t -> int
    end = struct
      type t = int
      let add x v = v + (abs x)
      let mult x v = v * (abs x)
      let read v = v
    end
  \end{ocaml}
\caption{\C{Counter} library in OCaml}
\label{fig:counter-adt}
\end{subfigure}
\begin{subfigure}[b]{0.56\textwidth}
  \begin{ocaml}
    module Counter: sig
      type t
      type eff 
      val add: int -> t-> eff
      val mult: int -> t -> eff
      val apply: eff -> t -> t
      val read: t -> int
    end = struct
      type t = int
      type eff = Add of int
      let add x v = Add (abs x)
      let mult x v = Add (v * (abs x - 1))
      let apply (Add x) v = x + v
      let read v = v
    end
  \end{ocaml}
\caption{\C{Counter} library re-engineered for effect-based replication}
\label{fig:counter-rdt}
\end{subfigure}
\end{figure}
The lack of composability is yet another important downside of this
approach.  Consider an application that uses two replicated counters,
$c_1$ and $c_2$, bound by the invariant $c_2 \ge c_1$, duly enforced
by the application when updating $c_2$ or $c_1$.  An execution may
nonetheless witness anamolous states that violate the invariant
because updates to $c_1$ and $c_2$ may be applied independently in any
order on any replica.  For example, if a replica increments both
counters, a \C{read} operation performed at another replica may
\begin{wrapfigure}{l}{.5\textwidth}
  \begin{center}
    \includegraphics[scale=0.4]{Figures/dali-counter}
  \end{center}
  \caption{\small A replicated counter can be expressed as a sequence
    of versions managed by logically-distributed concurrently
    executing threads.  Local versions produced by these threads
    periodically merge their state with one another.  In the figure,
    when the local version of Node 3, labeled 20, merges with the
    local version of Node 1, whose counter state is 15 at the time, a
    new counter value is produced.  This value takes into account the
    previous merged state (10) from which the two versions were
    derived to yield a new merged state of 25.  Observe that there is
    no explicit synchronization or coordination among versions.  In
    the figure, red circles represent counter state produced through
    merging, and blue circles represent state produced by applying
    counter operations on a node-local version.  }
\vspace*{-.85in}  
\end{wrapfigure}
witness the increment to $c_1$, but not $c_2$, thus violating the
invariant. The intention to atomically apply both effects cannot be
captured in the absence of external mechanisms to compose effects
(such as effect-based transactions~\cite{pldi15}).  Rather than
viewing a distributed computation in terms of \C{Add} effects produced
by operations, can we formulate a more declarative interpretation,
directly in terms of the counter value maintained by each node.  In
doing so, we observe that each replica essentially operates over its
own version of a counter.  Local operations on a replicated object can
be thought of as yielding new versions, thus producing a version tree,
with one branch for each replica.  Thus, each branch represents
different (immutable) versions maintained by different nodes, with the
state produced by a counter's operations on a node recorded along the
node's local branch.  Now, to generate a globally consistent view of a
counter, we only need to define a merge operation that explains how to
combine two local versions to produce a new version that reflects both
their states.

Framing replication as merging leads to a counter implementation that
bears strong similarity to the original sequential one:
  \begin{ocaml}
    module Replicated_Counter = struct
      include Counter
      type mergeable_t = t 
      let merge lca v1 v2 =
         lca + (v1 - lca) + (v2 - lca)
    end
  \end{ocaml}

The role of \C{lca} (least common ancestor) here captures salient
history - the state resulting from the merge of two versions derived
from the same ancestor state should not unwittingly duplicate the
contributions of the ancestor.  This interpretation of a replicated
datatype is thus given in terms of the evolution of a program state
impicitly associated with the different nodes that comprise a
distributed application with merge operations serving to communicate
and reconcile different local states.

In this paper, we develop a programming model that realizes a monadic
version control system centered around data, rather than file,
coherence.  The \name monad lets programmers write and compose
concurrent computations around multiple (implicit) versions of a
mergeable datatype, an ordinary ML datatype additionally equipped with
a \emph{merge} function responsible for deriving a consistent global
state from a collection of local versions of that state.  A
computation progresses by forking replicating new versions of existing
versions on geo-distributed nodes, creating new versions along a an
existing \emph{branch}, or by merging branch tips.

Our model allows ML programmers to get the benefits of realizing
highly available (low-latency) distributed computation, while
continuing to enjoy the comforts of high-level reasoning and the
familiarity of using standard libraries.  Notably, while \name's
programming model makes no explicit reference to any specific
operational manifestation of a distributed system (e.g., programmers
do not need to explicitly manage replicas), we demonstrate that it can
be nonetheless efficiently realized on existing geo-replicated
distributed systems.

Our contributions are summarized below:

\begin{itemize}
    \item We formally introduce the concept of a \emph{mergeable}
      datatype to admit high-level declarative reasoning about
      distributed computation and replicated state in ML programs.

    \item We describe the \name library that transparently adds
      persistence and replication features to a mergeable type. \name
      comes with a programming model that brings to bear the power and
      flexibility of a version control system to the administration of
      replicated data. \name hides the complexity of version control
      behind a monad, and exposes its functionality via a simple API
      that lets programmers define and compose distributed ML
      computations around such data.  Issues related to replication
      only manifest implicitly in the definition of a merge operation
      that defines coherence among different instances of replicated
      state.

    \item We present a formalization of the \name programming model,
      and establish the comditions under which eventual convergence of
      concurrent version state can be guaranteed.

    \item We describe an implementation of \name atop the Irmin
      persistent store~\cite{irmin}, a content-addressable storage
      library for OCaml, along with case studies and experimental
      results that justify its practical utility.
\end{itemize}

%% Persistence adds yet another dimension to the problem. Functional data
%% structures are often large in-memory linked structures that admit
%% functional updates by creating newer versions that share most of their
%% structure with previous versions. Sharing is the key to efficiency,
%% without which every functional update takes time that is at least
%% linear in the size of the data structure. Unfortunately, the
%% straightforward way to persist a data structure on disk in a
%% machine-agnostic format (e.g., JSON) requires serialization, which is
%% linear in the size of the data structure. An alternative is to
%% maintain a separate on-disk copy of the data and administer it via a
%% database system. The downside is that the programmer is now required
%% to reason in terms of a lower-level abstraction (e.g., a relational
%% data model) in order to maintain coherence between in-memory and
%% on-disk representations. Safety guarantees over the in-memory data do
%% not carry over to the disk, which introduces additional complications
%% and the possibility of new bugs. Thus, adding persistence to
%% applications while preserving high-level safety gurantees offered by
%% the language abstraction is a non-trivial problem.

%% However, as we shall see,
%% unconstrained branching may yield a branching structure that becomes
%% unmergeable, thus leading to \emph{stuck} states. An important
%% property of the \name library is that it guarantees progress even in
%% the presence of arbitrary merges provided the merge operation is
%% commutative with respect to its version state arguments.  Furthermore,
%% \name profitably exploits provenance information available via
%% branches to make application resilient to network partitions.  
%% In this paper, we define a semantics of mergeable datatypes that lets
%% any OCaml program be seamlessly deployed in a distributed environment
%% with implicit support for replicated and decentralized operation.  The
%% underlying foundation for merge actions has a natural category
%% theoretic interpretation that enables us to construct useful mergeable
%% variants for \emph{any} OCaml type whose states are commutative wtih
%% respect to the merge operation defie