### 2017-02-27 paper/evaluation.tex
edits

\section{Evaluation}
\label{sec:evaluation}

In this section, we present an evaluation of \name programming model with the
aim of assesing its practical utility for describing geo-distributed eventually
consistent applications. In particular, we

To this end, we have implemented a collaborative editing application that
simulates concurrent editing of the same document by several authors. This
application is not unlike Google Docs, but also differs from it by the fact
that we do not need a central server that coordinates the edits. Instead, edits
from each author is asynchronously sent to other authors. The shared document
itself is represented by a mergeable lists of strings
(Section~\ref{sec:model}), and hence, the remote edits can be correctly
reconciled into the local document. We resolve conflicts on concurrent
substitutions at the same position (such as two authors concurrently editing
\C{abc} to \C{xbc} and \C{ybc}, respectively) by choosing the character with
the smaller ASCII code.