### 2017-02-27 paper/rope-sig.tex

\begin{figure}

\begin{subfigure}{0.35\textwidth}
\begin{ocaml}
module type MRope = sig
  module A: sig
    type t
    val concat: t -> t -> t
    val split: t -> t*t
    val merge: t -> t -> t -> t
  end
  type t =
  | Leaf of A.t
  | Node of t * int * t
  ... 
  (* Standard rope functions *)
  val merge: t -> t -> t -> t 
end
\end{ocaml}
\caption{Signature of Mergeable Ropes}
\label{fig:mrope-sig}
\end{subfigure}
\begin{subfigure}{0.6\textwidth}
\begin{ocaml}
let rec flatten = function
  | Leaf x -> x
  | Node (l,_,r) -> A.concat (flatten l) (flatten r)

let rec merge old l r =
  if l = r then l else if old = l then r
  else if old = r then l else merge_rec old l r

and merge_rec old l r = match (old,l,r) with
  | Leaf _, _, _ | _, Leaf _, _ 
  | _, _, Leaf _ -> 
      rope_of @@ A.merge (flatten old) 
                    (flatten l) (flatten r)
  | Node (oldl,_,oldr), Node (ll,_,lr), 
    Node (rl,_,rr) ->
      let newl = merge oldl ll rl in
      let newr = merge oldr lr rr in
        Node (newl, create_index newl, newr)
\end{ocaml}
\caption{Signature of Mergeable Lists}
\label{fig:mrope-merge}
\end{subfigure}

\caption{Mergeable List Implementation}
\label{fig:rope}
\end{figure}