### 2017-02-28 paper/evaluation.tex
edits

\section{Evaluation}
\label{sec:evaluation}

In this section, we present an evaluation of \name programming model
with the aim of assessing its practical utility for describing
geo-distributed eventually consistent applications. In particular, we
are interested in the horizontal scalability of applications built
using \name. Typical distributed eventually consistent data stores use
custom synchronization and dissemination protocols for transferring
updates between replicas. In contrast, \name is realized over Irmin~\cite{Irmin}
which uses Git transfer protocol~\cite{..}, a protocol not designed
with distributed data stores in mind. Hence, we also evaluate the
performance of synchronization and merging.

\subsection{Benchmark: Collaborative editing}

To evaluate the effectiveness of our model, we have implemented a
collaborative editing application that simulates concurrent editing of
the same document by several authors. This application is not unlike
Google Docs, but differs from it by the fact that we do not need a
central server that coordinates the edits.  Instead, edits from each
author is asynchronously sent to other authors. The shared document
itself is represented by a mergeable rope of characters
(Section~\ref{sec:model}), and hence, the remote edits can be
correctly reconciled into each local document. We resolve conflicts on
concurrent substitutions at the same position (such as two authors
concurrently editing \C{"abc"} to \C{"xbc"} and \C{"ybc"},
respectively) by choosing the character with the smaller ASCII
code. For the benchmarks, the initial document that we use has 1576
characters. The workload consists of 4000 edit operations at random
indices with 85\% insertions and 15\% deletions.

\subsection{Experimental setup}

\begin{figure}[t]
  \centering
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[width=0.9\textwidth]{Figures/cluster.pdf}
		\caption{Our experimental configuration consists of an
                  8-node ring cluster on executing on a Google Cloud
                  Platform. US-C, EU-W, ASIA-NE and ASIA-E are
                  availability zones us-central1-b, europe-west1-b,
                  asia-northeast1-b and asia-east1-b
                  respectively. Edge labels are inter-node latencies
                  in milliseconds.}
		\label{fig:cluster}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[width=\textwidth]{Graphs/scalability.pdf}
		\caption{Scalability: Overall throughput of the cluster and latency of each
			operation.}
		\label{grf:scalability}
	\end{subfigure}
	\par\bigskip
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[width=\textwidth]{Graphs/synchronization.pdf}
		\caption{Synchronization: Overall throughput of the cluster while actively
			synchronizing with the successor node.}
		\label{grf:synchronization}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[width=\textwidth]{Graphs/merge.pdf}
		\caption{Merge performance: Cost of merging concurrent operations across
		nodes.}
		\label{grf:merge}
	\end{subfigure}
	\caption{\name performance evaluation.}
  \label{grf:evaluation}
\end{figure}

For our experiments, we instantiate an 8-node geo-distributed cluster
using the Google Cloud Platform~\cite{gcp}. Each node is an
\C{n1-standard-4} instance with 2 virtual CPUs and 7.5 GB of
memory. The nodes are arranged in a ring as shown in
Figure~\ref{fig:cluster} and synchronize (push and pull updates) when
necessary with its successor. While \name does not impose any
restrictions on synchronizing with other nodes, we choose to only
synchronize with the successor since we are guaranteed eventual
convergence with a ring cluster. If the successor becomes unavailable,
the node synchronizes with other available nodes. Operations are
generated and applied locally on each node before they are propagated
eventually to other nodes with the help of a background thread that
synchronizes with its successor every second.

\subsection{Results}

We evaluate the scalability of concurrent editing application by increasing the
cluster size from 1 to 8 (the 4 node ring cluster consists of nodes numbered 0
to 3), with each node performing concurrent edits to the same document. In each
case, we measure the overall cluster throughput and latency of each operation.
The results are presented in Figure~\ref{grf:scalability}. The results show
that the cluster throughput increases linearly with the number of concurrent
editors, while the latency for each operation remains the same. This is because
each operation is performed locally and does not require synchronization with
other nodes. The nodes remain available to accept requests even if the node
gets disconnected. Since the document type is mergeable, eventually when the
node comes back online, the updates are synchronized with the cluster.

While updates are propagated asynchronously, we also measured the impact of
propagating updates synchronously by forcing synchronization every few
operations. This corresponds to a partially synchronous system, which is also
unavailable; if the cluster is unreachable, local operations will not be
accepted. While this is not a realistic deployment, the results help us
understand the overhead of synchronization. The results are shown in
Figure~\ref{grf:synchronization}. The result show that frequent blocking
synchronization is prohibitively expensive with the throughput dropping to 3
ops/s for synchronizing after every operation. \name uses Git transfer protocol
for synchronization which involves multiple round trips between the nodes for
synchronization. But as the synchronizations get less frequent, we can see that
the throughput asymptotically approaches the asynchronous case
(Figure~\ref{grf:scalability}).

Finally, we evaluate the performance of merge functions. For this experiment,
we consider the nodes 1 (\C{us-central1-b}) and 2 (\C{europe-west1-b}). Both
nodes start with the initial document and perform $n$ concurrent edits
without synchronization. After this, the nodes synchronize pushing the local
edits and pulling the remote edits, and merge to produce the final document. In
this case, we indicate the distance of synchronization as $2n$.
Figure~\ref{grf:merge} presents the time share between network and computation
for performing 3-way merge as we increase the operation distance between the
remote branches. We see that the merge time is dominated by the network
overhead. This is because of the transfer protocol involving multiple
round trips and the need to transfer objects not only for application objects
but also Git internal objects for history, directory structure and branches. We
observe that we can improve the network performance by implementing a streaming
protocol that eagerly transfers objects.

More importantly, the cost of computation grows sub-linearly. This is
because of efficient merge function on ropes (Sec.~\ref{sec:ropes})
enabled by Irmin's content-addressed storage backend, and more
expensive operational transformation of merging concurrent edits to
leaf nodes (Sec.~\ref{sec:lists}) is invoked rarely on small. On
average, we measured that with a distance of 256, for merging the two
documents with 1743 characters, 50 list merges were performed on
strings of average length 13. A purely list based approach would
require a polynomial time edit sequence algorithm to be applied on the
entire document twice. Compared to this, the rope implementation has
much better algorithmic complexity.  While not depicted here, we
measured that by representing the entire document using a mergeable
list type, in addition to slowing down each edit operation, the merge
at a distance of 256 was 2.3 times slower than the rope
implementation.  We conclude that versioning and merging as found in
\name\ can yield substantial throughput and latency gains with low
computational cost on modern eventually-consistent cloud systems.