### 2017-02-28 paper/appendix.tex
Committing meta.

\section{Appendix}
\input{list-code}
\begin{lemma} [\bfseries Unique External Locus]
Every version $v$ (except the \C{INIT} version) in the history $H$
generated by the rules in Fig.~\ref{fig:opsem} has a unique external
locus.
\end{lemma}
\begin{proof}
Follows from the definitions~\ref{def:external-locus},
\ref{mergeability}, and the \rulelabel{E-Pull-Wait} rule.
\end{proof}

\begin{lemma} [\bfseries Unique LCA]
Every pair of branches in a legal history $H$, if mergeable as per
Def.~\ref{fig:mergeability}, have a unique least common ancestor. 
\end{lemma}
\begin{proof}
By construction. If $b_j$ is mergeable into $b_i$, the locus of
$b_i$ is the least common ancestor of $b_i$ and $b_j$
\end{proof}

\begin{lemma}
Every pair of 
\end{lemma}
% If $b_j$ is not mergeble into $b_i$, it could be because:
% 1. $b_j$ included some internal nodes of $b_i$ that are not in
% $L(b_i)$, then simply merge $b(L(b_i))$ into $b_i$, and then that back
% into $b(L(b_i))$. Now the problem reduces to merging $b(L(b_i))$
% 2. $If