### 2017-02-28 paper/implementation.tex
edits

\section{Distributed Instantiation}
\label{sec:implementation}

We describe the realization of \name programming model on top of a persistent
distributed store and the language extensions to enable and enforce the
constraints of the \name.

\name programming model is realized on top of Irmin~\cite{irmin}, an OCaml
library database implementation that is part of the MirageOS
project~\cite{mirage}. Irmin provides a persistent multi-versioned store with
with a content-addressable heap abstraction. Simply put, content-addressability
means that the address of a data block is determined by its content. If the
content changes, then so does the address. Old content continues to be
available the old address. Content-addressability also results in constant time
structural equality checks, which we exploit in our mergeable rope
implementation (Section~\ref{sec:ropes}).

Irmin provides support for distribution, fault-tolerance and concurrency
control by incorporating Git distributed version control~\cite{git} protocol
over its object model. Indeed, Irmin is fully compatible with the Git command
line tools. Distributed replicas in \name are created by cloning a \name
repository. Due to \name's mergeable type, each replica can operate completely
independently, accepting client requests, even when disconnected from other
replicas, resulting in a fully available distributed database system.

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
	\caption{Network partitions may leave branches $b_1$ and $b_4$ in one
		partition (red background), and branches $b_2$ and $b_3$ in the other (yellow
		background). Since $b_2$ and $b_3$ cannot merge unless $b_1$ and $b_4$ merge
		(Fig.~\ref{fig:legal-extensions}), they cannot communicate among themselves
		even though they are in the same network partition.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked within the notion of branches,
allowing the programmer to explicitly merge the branches on demand. \name
realizes the concurrent programming model on top of the notion of branches;
each replica operates on its own branch, and thus is isolated from actions on
other replicas. In presence of network partitions, the nodes in one partition
may not receive updates from the other partition, hence merges cannot happen
across a partition. This can be a problem if a pair of branches in one
partition (the current partition) have multiple LCAs in the other partition
(the remote partition), hence cannot merge. Fig.~\ref{fig:partitions}
illustrate this problem for the branching structure of
Fig.~\ref{fig:external-lcas}.

Fortunately, the access to full (causal) branching history assured by Irmin at
every node comes to the rescue. Relying on the history, the current partition
can fork-off (\C{fork}) new branches that start where the remote branches left,
and map them to new (virtual) nodes that emulate the nodes in the remote
partition. For the example in Fig.~\ref{fig:partitions}, the partition of $b_1$
and $b_4$ can fork new branches $b_2'$ and $b_3'$ from $b_2$ and $b_3$,
respectively, and resume the activity as if the partition never happened. After
merging the first versions on $b_2'$ and $b_3'$, branches $b_1$ and $b_2$ will
be in the same state as before (when the partition happened), except that the
branches are now mergeable because $b_2'$ and $b_3'$ can merge. The ability to
track full provenance information is thus crucial for \name to overcome network
partitions, making it an appropriate programming model for highly-available
replicated data types. \name monad conceals the branching structure, but also
transparently performs the necessary merges to obtain a history graph with a
unique lowest common ancestor.

Importantly, Irmin supports user-defined three-way merge functions for
reconciling concurrent operations. While Irmin's merge functions are defined
over objects on Irmin's content-addressable heap, \name's merge functions are
defined over OCaml types. We address this representational mis-match with the
help of OCaml's PPX metaprogramming support~\cite{ppx} to derive bi-directional
transformations between objects on OCaml and Irmin heaps. We also derive the
various serialization functions required Irmin. This address the representation
mismatch between the application and storage layers. Synchronization between
replicas is performed using Git's notion of pushing and pulling updates from
remotes over the Git transfer protocol~\cite{git-tp}. As a result, we reap the
benefits of efficient delta-transfer (only missing objects are transferred
between replicas), compression and end-to-end encrypted communication between
the replicas.