### 2017-02-28 paper/implementation.tex
Edits.

\section{Distributed Instantiation}
\label{sec:implementation}

We describe the realization of \name on top of a persistent
distributed store and the language extensions to enable and enforce
the constraints of the \name.

\name programming model is realized on top of Irmin~\cite{irmin}, an OCaml
library database implementation that is part of the MirageOS
project~\cite{mirage}. Irmin provides a persistent multi-versioned store with
with a content-addressable heap abstraction. Simply put, content-addressability
means that the address of a data block is determined by its content. If the
content changes, then so does the address. Old content continues to be
available the old address. Content-addressability also results in constant time
structural equality checks, which we exploit in our mergeable rope
implementation (Section~\ref{sec:ropes}).

Irmin provides support for distribution, fault-tolerance and concurrency
control by incorporating Git distributed version control~\cite{git} protocol
over its object model. Indeed, Irmin is fully compatible with the Git command
line tools. Distributed replicas in \name are created by cloning a \name
repository. Due to \name's mergeable type, each replica can operate completely
independently, accepting client requests, even when disconnected from other
replicas, resulting in a fully available distributed database system.

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
  \caption{Network partitions may leave branches $b_1$ and $b_4$ in
  one partition (red background), and branches $b_2$ and $b_3$ in the
  other (yellow background). Access to full history on both sides of
  the partition lets both sides make progress.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked within the notion of branches,
allowing the programmer to explicitly merge the branches on demand. \name
realizes the concurrent programming model on top of the notion of branches;
each replica operates on its own branch, and thus is isolated from actions on
other replicas. In presence of network partitions, the nodes in one partition
may not receive updates from the other partition, hence merges cannot
happen across a partition. This can be a problem if a pair of branches
in one partition (the current partition) have their loci
(Sec.~\ref{sec:meta}) in the other partition (the remote partition),
hence cannot merge. Fig.~\ref{fig:partitions} illustrate this problem
for the branching structure of
Fig.~\ref{fig:external-lcas}.

Fortunately, the access to full (causal) branching history assured by
Irmin at every node comes to the rescue. Relying on the history, the
current partition can fork-off (\C{fork}) new branches that start
where the remote branches left, and map them to new (virtual) nodes
that emulate the nodes in the remote partition. For the example in
Fig.~\ref{fig:partitions}, the partition of $b_1$ and $b_4$ can fork
new branches $b_2'$ and $b_3'$ from $b_2$ and $b_3$, respectively, and
resume the activity as if the partition never happened. The ability to
track full provenance information is thus crucial for \name to
overcome network partitions, making it an appropriate programming
model for highly-available replicated data types. \name monad conceals
the branching structure, but also transparently performs the necessary
merges to obtain a history graph with a unique lowest common ancestor.

Importantly, Irmin supports user-defined three-way merge functions for
reconciling concurrent operations. While Irmin's merge functions are
defined over objects on Irmin's content-addressable heap, \name's
merge functions are defined over OCaml types. We address this
representational mismatch with the help of OCaml's PPX metaprogramming
support~\cite{ppx} to derive bi-directional transformations between
objects on OCaml and Irmin heaps. We also derive the various
serialization functions required by Irmin\footnote{
  The Since \nameMonad monad relies on Irmin for realizing the branch
  structure, it is in fact parameterized over a mergeable type that
  implements all these functions. Thus the monad is functorized over
  the mergeable type, but not polymorphic. The convenience of type
  parameter inference that polymorphism offers can however be obtained
  for module parameters using OCaml extensioms, such as Modular
  Implicits~\cite{implicits}.
}.
Synchronization between replicas is performed using Git's notion of
pushing and pulling updates from remotes over the Git transfer
protocol~\cite{git-tp}. As a result, we reap the benefits of efficient
delta-transfer (only missing objects are transferred between
replicas), compression and end-to-end encrypted communication between
the replicas.