### 2017-02-28 paper/implementation.tex

\section{Distributed Instantiation}
\label{sec:implementation}

We describe the realization of \name on top of a persistent
distributed store and the language extensions to enable and enforce
the constraints of the \name.

The \name programming model is realized on top of Irmin~\cite{irmin},
an OCaml library database implementation that is part of the MirageOS
project~\cite{mirage}. Irmin provides a persistent multi-versioned
store with a content-addressable heap abstraction. Simply put,
content-addressability means that the address of a data block is
determined by its content. If the content changes, then so does the
address. Old content continues to be available from the old
address. Content-addressability also results in constant time
structural equality checks, which we exploit in our mergeable rope
implementation (Section~\ref{sec:ropes}), among others.

Irmin provides support for distribution, fault-tolerance and
concurrency control by incorporating the Git distributed version
control~\cite{git} protocol over its object model. Indeed, Irmin is
fully compatible with Git command line tools. Distributed replicas in
\name are created by cloning a \name repository. Due to \name's
support for mergeable types, each replica can operate completely
independently, accepting client requests, even when disconnected from
other replicas, resulting in a highly available distributed system.

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
  \caption{Network partitions may leave branches $b_1$ and $b_4$ in
  one partition (red background), and branches $b_2$ and $b_3$ in the
  other (yellow background). Access to full history on both sides of
  the partition lets both sides make progress.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked using the notion of
branches, allowing the programmer to explicitly merge branches on
demand. \name's concurrency and distribution model is also realized in
terms of branches; each replica operates on its own branch, and thus
is isolated from the actions on other replicas. In the presence of
network partitions, the nodes in one partition may not receive updates
from the other partition, preventing merges from happening across a
partition. This can be a problem if a pair of branches in one
partition (the current partition) have multiple LCAs in the other
partition (the remote partition), and hence cannot
merge. Fig.~\ref{fig:partitions} illustrate this problem for the
branching structure of Fig.~\ref{fig:external-lcas}.

Fortunately, access to full (causal) branching history assured by
Irmin at every node comes to the rescue. Relying on the history, the
current partition can fork-off (\C{fork}) new branches that start
where the remote branches left, and map them to new (virtual) nodes
that emulate the nodes in the remote partition. For the example in
Fig.~\ref{fig:partitions}, the partition of $b_1$ and $b_4$ can fork
new branches $b_2'$ and $b_3'$ from $b_2$ and $b_3$, respectively, and
resume the activity as if the partition never happened. After merging
the first versions on $b_2'$ and $b_3'$, branches $b_1$ and $b_2$ will
be in the same state as before (when the partition happened), except
that the branches are now mergeable because $b_2'$ and $b_3'$ can
merge. The ability to track full provenance information is thus
crucial for \name to overcome network partitions, making it an
appropriate programming model for highly-available replicated data
types. The \nameMonad conceals branching structure, but also
transparently performs the necessary merges to obtain a history graph
with a unique lowest common ancestor.

Importantly, Irmin supports user-defined three-way merge functions for
reconciling concurrent operations. While Irmin's merge functions are
defined over objects on Irmin's content-addressable heap, \name's
merge functions are defined over OCaml types. We address this
representational mismatch with the help of OCaml's PPX metaprogramming
support~\cite{ppx} to derive bi-directional transformations between
objects on OCaml and Irmin heaps. We also derive the various
serialization functions required by Irmin\footnote{
  The Since \nameMonad monad relies on Irmin for realizing the branch
  structure, it is in fact parameterized over a mergeable type that
  implements all these functions. Thus the monad is functorized over
  the mergeable type, but not polymorphic. The convenience of type
  parameter inference that polymorphism offers can however be obtained
  for module parameters using OCaml extensioms, such as Modular
  Implicits~\cite{implicits}.
}.
Synchronization between replicas is performed using Git's notion of
pushing and pulling updates from remotes over the Git transfer
protocol~\cite{git-tp}. As a result, we reap the benefits of efficient
delta-transfer (only missing objects are transferred between
replicas), compression and end-to-end encrypted communication between
he replicas.