### 2017-02-28 paper/conc.tex

\section{Conclusions}
\label{sec:conclusions}

Replication is a critical feature in geo-distributed systems used to
improve latency and availability.  Existing approaches to dealing with
replicated data often involve complex and subtle code restructuring,
and the use of specialized datatypes that interact poorly with other
language structures.  \name\ is a programming model that extends ML
with implicit support for replication that overcomes these concerns.
Its notion of concurrency and synchronization is encapsulated within a
monad that considers distrbuted computation in terms of typed versions
of replicated state.  Notably, any ML data structure can participate
in a distributed computation using this monad by providing a merge
function that dictates how different versions of that structure can be
reconciled.  Our versioning semantics enjoys strong consistency and
progress guarantees that make it useful in a fully decentralized (and
unreliable) cloud environment.  Furthermore, our experimental results
demonstrate that using such high-level abstractions need not come at
the expense of efficient implementations.