### 2017-03-15 paper/examples.tex
Fixed many errors.

\section{Mergeable Types}
\label{sec:mergeable_types}

\input{rope-sig}

The \drawsome example demonstrates \name's utility in building
concurrent applications with conceptual state sharing, which makes
them suitable for transparent deployment in a distributed setting. The
model allows concurrent computations to be composed around any
mergeable type. We now present a series of examples of mergeable types
built either by defining merge functions for ML data types, or by
composing existing mergeable types.

\subsection{Ropes}
\label{sec:ropes}

A rope is a data structure that facilitates efficient implementations
of operations such as concatenation, splitting, lookup and insertion
at a random position etc., for a list of values. The values are
usually characters, their lists being strings. A rope is essentially a
binary tree, whose leaf nodes store substrings of a larger string such
that their left-to-right (in-order) concatenation yields the larger
string. The idea is to facilitate efficient lookups and insertions at
random positions in the string by storing position indexes in the
internal nodes of the tree.

Fig.~\ref{fig:mrope-sig} shows a rope interface in OCaml. The type of
the list of values is left abstract (\C{A.t}); it can be instantiated
with any module that supports \C{concat} and \C{split} operations. The
interface shown in Fig.~\ref{fig:mrope-sig} is that of a mergeable
rope (\C{MRope}), whose \C{merge} implementation is shown in
Fig.~\ref{fig:mrope-merge}. The function returns one of its versions
(\C{l} or \C{r}), if the other one is the same as the common ancestor.
Otherwise, it calls \C{merge\_rec}, which recursively compares the
structures of merging ropes, and their common ancestor.  The function
recurses until one of the ropes differs from the other two, or all of
them are leaf nodes, at which point it relies on the (abstract) list
merge function (\C{A.merge}) to merge the lists represented by the
ropes. \C{merge\_rec} uses auxiliary functions, \C{mk\_rope} and
\C{create\_index} (definitions not shown), respectively, to construct
a rope given a list (\C{A.t}), and to compute the index at a node
given its left sub-tree.

A notable feature of the mergeable rope implementation, which is also
a characteristic of mergeable types in general, is compositionality. A
rope is a polymorphic mergeable data type that composes mergeability
over its polymorphic constituents.  Compositionality is concretized by
the \C{MRope} signature, which requires the type of items in the list
(\C{A.t}) to also be mergeable (\C{A.merge}).

\subsection{Lists}
\label{sec:lists}

\input{list-sig}

As an abstract data type, \C{list} supports three operations:
\C{insert x i l}, to insert an element \C{x} at position \C{i} in the
list \C{l}, \C{delete i l}, to delete the element at \C{i}'th position
in \C{l}, and \C{subst i x l}, to substitute the element at the
\C{i}'th position with \C{x} in \C{l}.  Lists of mergeable items are
mergeable if they preserve the following properties:
\begin{itemize}
  \item Elements present in both the merging lists will be present in
  the final list. No element is deleted unless it is deleted in at
  least one list.
  \item An element deleted in one list will not reappear in the merged
  list (unlike, for example, Amazon's replicated shopping cart~\cite{Dynamo}).
  \item The order of elements is retained. That is, if $x$ occurs
  before $y$ in one of the lists, and if both elements are present in
  the merged list, then $x$ occurs before $y$ in the merged list.
\item If an element $y$ is substituted for an element $x$ in one list,
  then $y$ is also substituted for $x$ in the merged list (i.e., $x$
  is absent and $y$ is present), unless $x$ is deleted or substituted
  in the other list. In the former case, deletion wins, and neither
  $x$ nor $y$ occur in the merged list. In the latter case, if $x$ is
  substituted by a different element $z$ in the other list, then the
  merged list substitutes $x$ by a merge of $y$ and $z$, defined as
  per the merge semantics of the list element type.
\end{itemize}

\noindent The merge operation for lists is composed of two separate functions,
\C{edit\_seq} and \C{op\_transform}. Both these functions have been
implemented in less than 50 lines of OCaml.  The functions are described below.
\begin{itemize}
	\item \C{edit\_seq} takes a pair of lists, \C{v} and $\C{v'}$, and computes
		the shortest sequence of list operations that need to be applied on \C{v}
		to obtain \C{v'}. Such a sequence is called an \emph{edit sequence}. The
		length of the sequence corresponds to the standard notion of the \emph{edit
		distance} between the two lists, which can be computed in polynomial
		time~\cite{wagner-fischer}.  \C{edit\_seq} modifies one such algorithm to
		also compute the edit sequence (an \C{edit list}). The signature
		specified in Fig.~\ref{fig:mlist} represents edits using the type
		\C{edit}.  Constructors \C{I}, \C{D}, and \C{S} stand for \C{insert},
		\C{delete}, and \C{subst}, respectively.\footnote{ Note that \C{I}, \C{D}
		and \C{S} are \emph{not} effects; no list function generates them, and
		they are not exchanged between concurrent threads. They are simply tags
		used by the list merge operation for convenience.  Indeed, to better
		appreciate the virtues of mergeable types, we encourage the reader to
		think about how one might implement a replicated list with this
		functionality using an explicit effect representation as described in
		Sec.~\ref{sec:intro}.}  The subst constructor also carries the \C{A.t}
		element that was substituted. The element serves as the \C{lca} argument to
		\C{A.merge} in case of concurrent conflicting substitutions.
		Fig.~\ref{fig:list-eg} illustrates edit sequences for a sample list.  The
		sequence \C{[I(c,0); S(3,c,s)]} maps the list \C{[a;b;c]} to \C{[c;a;b;s]}
		because \C{subst 3 s (insert c 0 [a;b;c]) = [c;a;b;s]}.
	\item  \C{op\_transform} takes a pair of edit sequences, $s_1$ and $s_2$,
		that map a list $v$ to two different lists, $v_1$ and $v_2$ (e.g.,
		Fig.~\ref{fig:list-eg}), and transforms $s_1$ to $s_1'$ such that $s_1'$
		has the same effect on $v_2$ as $s_1$ had on $v$.  For instance, in
		Fig.~\ref{fig:list-eg}, let $s_1$ be the edit sequence \C{[D(1); S(1,c,d)]},
		which maps the list \C{[a;b;c]} ($v$) to \C{[a;d]} ($v_1$). The \C{D} edit
		deletes the first element of $v$. However, if $s_1$ is to be applied on the
		list \C{[c;a;b;s]} ($v_2$), which has a new element at position 0, the
		\C{D} edit must delete the second element from $v_2$, if it were to
    have the same effect as it did on $v$. We say \C{D(1)} in $s_1$
    has to be transformed w.r.t the sequence $s_2 =
    \C{[I(c,0);S(3,c,s)]}$.  The function \C{op\_transform} computes
    such \emph{operational transformations} for an edit sequence $s_1$
    w.r.t another edit sequence $s_2$. A notable transformation rule
    is for a substitution w.r.t another substitution at the same
    position. Since the substituting items could be different, the
    function relies on the item merge function (\C{A.merge}) to merge
    them into a single item. For instance, in Fig.~\ref{fig:list-eg},
    there are simultaneous substitutions to the item \C{c} in $s_1$
    and $s_2$; while $s_1$ substitutes it with \C{d}, $s_2$ does it
    with \C{s}. The merged list therefore substitutes \C{c} with the
    merge of \C{d} and \C{s}, defined as per the merge semantics of
    \C{A.t}.
\end{itemize}
The \C{MList.merge} can now be defined as a function that first
computes edit sequences, $s_1$ and $s_2$, for the two concurrent
lists, $v_1$ and $v_2$ w.r.t their common ancestor $v$, transforms one
of them (say, $s_1$) w.r.t $s_2$ to obtain $s_1'$, and finally applies
(the operations defined by) $s_1'$ to $v_2$ to obtain the merged list.

\subsection{Lists of Integers and Quantities}.

A mergeable list of integers can be composed using mergeable lists and
mergeable integers.  Concretely, it is a composition of \C{MList} and \C{Int}
types (i.e., \C{MList(Int)}), where \C{Int.merge} is the same as
\C{Counter.merge} described in the introduction.  An alternative instantiation
of \C{MList} is for a quantity type (\C{Qty}) that supports \C{add} and
\C{subtract} operations, and defines its \C{merge} operation a little
differently:
\begin{ocaml}
let merge lca v1 v2 = if v1 = lca then v2
    else (if v2 = lca then v1 else max v1 v2)
\end{ocaml}
\begin{wrapfigure}{l}{.45\textwidth}
  \centering
  \includegraphics[scale=0.4]{Figures/list-eg}
  \caption{Lists of mergeable values are mergeable. }
  \label{fig:list-eg}
  \vspace*{-.45in}
\end{wrapfigure}
\C{MList(Qty)}, like \C{MList(Int)}, is a list of integers, and has the same
semantics under a sequential execution. Their differences however manifest
under a concurrent execution, where both adopt different methods to reconcile
concurrent updates.

If efficient insertions and lookups (by position) are desired over a
large list of integers, while admitting concurrent updates, we require
a mergeable rope of integers, which can be obtained by simply
composing \C{MList(Int)} with \C{MRope} (i.e., \C{MRope(MList(Int))}).

% The mergeable list data type demonstrates a common pattern of writing
% \C{merge} functions applicable to a range of functional data
% structures. For instance, a binary tree (with rotations) can be made
% mergeable in the same way as list, i.e., by computing edit sequences
% (\C{edit\_seq}) and then their transformations
% (\C{op\_transform})\footnote{
%   Like Wagner-Fischer, established algorithms exist for computing tree
%   edit distances~\cite{tree-diff}. An alternative naive method to
%   merge a pair of trees is by merging the lists obtained from their
%   in-order traversal, followed by reconstructing the tree.
% }.

\subsection{Shopping List}

\begin{figure}
\centering
\begin{tabular}{l||l}
\begin{ocaml}
let alice_f : L.t unit t =
  fork bob_f >>= fun () -> (* Invite Bob *)
  get () >>= fun c0 ->
  let qty = L.assoc "eggs" in
  let c0' = L.subst 1 ("eggs",qty+1) c0 in
  sync () ~v:c0' >>=
  fun c1 -> return ()
\end{ocaml}
&
\begin{ocaml}
let bob_f : L.t unit t =
  get () >>= fun c0 ->
  let c0' = L.insert ("candy",Qty.of_int 1) 1 @@
            L.delete @@ L.lookup "milk" in
  sync () ~v:c0' >>=
  fun c1 -> return ()
\end{ocaml}
\\
\end{tabular}
\caption{Bob and Alice collaboratively build a grocery shopping list}
\label{fig:shopping-list-code}
\end{figure}

A non-trivial demonstration illustrating the compositional power of
mergeable types is a shopping list application that allows its users
to collaboratively build a shopping list (listing items in the
decreasing order of their priority), by adding new items with a
specified quantity, deleting existing items, and increasing or
decreasing their quantity. An item (\C{Item.t}) is represented as a
tuple of its name (a string) and the quantity (A \C{Qty.t}).  The
merge function for \C{Item.t} merges the quantities of items (using
\C{Qty.merge}) having the same names. Items with different names are
considered distinct. A shopping list is a mergeable list of items,
i.e., \C{MList(Item)} (although a grocery shopping list may not be
large enough to justify using ropes, \C{MRope(MList(Item))} is
certainly allowed).  \C{MList(Item).merge} automatically lifts the
merge semantics for items to a merge semantics for lists of items.
Fig.~\ref{fig:shopping-list} illustrates its behavior.  Here, two
users, Alice and Bob, concurrently edit a shopping list whose initial
contents are \C{milk} and \C{eggs}. While both simultaneously update
the quantity of eggs, Bob also removes \C{milk} and inserts \C{candy}.
Fig.~\ref{fig:shopping-list-code} shows the \name code for two
concurrent sessions. The merged shopping list contains \C{eggs} and
\C{candy}, where the quantity of eggs is obtained by merging Alice's
\begin{wrapfigure}{r}{.35\textwidth}
  \centering
  \includegraphics[width=0.25\textwidth]{Figures/shopping-list}
  \caption{Merging concurrent shopping lists}
  \label{fig:shopping-list}
  \vspace*{.25in}
\end{wrapfigure}
and Bob's quantity as per the definition of \C{Qty.merge}.

The examples described in this section make certain assumptions about
the concurrency model and the underlying system, such as the existence
of a single lowest common ancestor (LCA) for any pair of versions, the
ability to access a previous version on any branch, and mergeability
of any two concurrent versions. Enforcing these guarantees in a fully
decentralized distributed setting requires addressing non-trivial
theoretical challenges. These, and other details that contribute to
the practicality of our model, such as containing the complexity of
merges, are discussed in subsequent sections.